Transition path sampling based calculations of free energies for enzymatic reactions: the case of human methionine adenosyl transferase and plasmodium vivax adenosine deaminase

### 2022-06-08 tex/resubmission1/manuscript.tex
added some references

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  Transition path sampling (TPS) is widely used for the
  calculations of reaction rates, transition state structures and reaction 
  coordinates of condensed phase systems. Here we discuss a scheme 
  for the calculation of free energies using the ensemble of TPS reactive 
  trajectories in combination with a window based sampling technique 
  for enzyme catalyzed reactions. We calculate the free energy profiles of 
  the reactions catalyzed by the human methionine S-adenosyltransferase (MAT2A) 
  enzyme and the plasmodium vivax adenosine deaminase ($pv$ADA) enzyme to assess 
  the accuracy of this method. MAT2A catalyzes the formation of S-adenosine-L-methionine
  following a S$_{\text{N}}2$ mechanism and using our method we estimate
  the free energy barrier for this reaction to 
  be $16\;\text{Kcal}\;\text{mol}^{-1}$ which
  is in excellent agreement with the experimentally measured activation 
  energy of $17.27\;\text{Kcal}\;\text{mol}^{-1}$. Furthermore, for the 
  $pv$ADA enzyme catalyzed reaction we estimate a free energy barrier of 
  $23\;\text{Kcal}\;\text{mol}^{-1}$ and the calculated free energy profile is
  similar to that predicted from experimental observations.
  Calculating free energies employing our simple method within TPS 
  provides significant advantages over such methods as umbrella sampling 
  since it is free from any applied external bias, accurate compared to 
  experimental measurements and has a reasonable computational cost.    
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Molecular dynamics (MD) simulations are increasingly being
used to calculate quantitative estimates of experimentally measurable 
kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
\cite{Karplus02NatStructMolBiol9p646,
Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
This has helped in proposing mechanisms and explanations for 
enzyme activity, fast high throughput screening of thousands of inhibitor 
molecules in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
directed evolution for designing enzymes with enhanced catalytic 
activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
Straightforward MD simulations adequately sample 
the long-lived stable reactant and product states 
but the high energy barrier-crossing events that occur rarely and at long 
time scales are not easily accessible using this method alone.
To access such states at a higher frequency using MD simulations 
several enhanced sampling techniques 
have been developed in the past few decades. These include 
umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
steered MD, \cite{Park04JChemPhys120p5946}
milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,
Dellago98JChemPhys108p1964,Bolhuis21AdvTheory4p2000237}

TPS is an attractive method since it samples reactive trajectories
in which the rare barrier-crossing events are guaranteed to occur and
it does not require a priori knowledge of the reaction coordinate which can be 
complex for large biological systems.
TPS is based on Monte Carlo importance sampling in the 
space of trajectories. It uses the shooting and shifting algorithm 
to collect an ensemble of reactive pathways that are real dynamical 
trajectories generated without the application of any bias forces. \cite{dellago02AdvChemPhys123} 
Therefore the TPS ensemble can be used to obtain 
kinetic information such as the rate of the reaction as well as to elucidate 
the transition state, reaction coordinate and the reaction mechanism. 
TPS has been successfully applied to study problems such as enzyme catalysis, 
\cite{Antoniou01JPhysChemB105p5553,Hay12NatChem4p161,Schwartz09NatChemBiol5p551} 
and protein folding. \cite{Bolhuis03ProcNatlAcadSci100p12129,Juraszek12ChemPhys396p30}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  Transition path sampling (TPS) is widely used for the
  calculations of reaction rates, transition state structures and reaction 
  coordinates of condensed phase systems. Here we discuss a scheme 
  for the calculation of free energies using the ensemble of TPS reactive 
  trajectories in combination with a window based sampling technique 
  for enzyme catalyzed reactions. We calculate the free energy profiles of 
  the reactions catalyzed by the human methionine S-adenosyltransferase (MAT2A) 
  enzyme and the plasmodium vivax adenosine deaminase ($pv$ADA) enzyme to assess 
  the accuracy of this method. MAT2A catalyzes the formation of S-adenosine-L-methionine
  following a S$_{\text{N}}2$ mechanism and using our method we estimate
  the free energy barrier for this reaction to 
  be $16\;\text{Kcal}\;\text{mol}^{-1}$ which
  is in excellent agreement with the experimentally measured activation 
  energy of $17.27\;\text{Kcal}\;\text{mol}^{-1}$. Furthermore, for the 
  $pv$ADA enzyme catalyzed reaction we estimate a free energy barrier of 
  $23\;\text{Kcal}\;\text{mol}^{-1}$ and the calculated free energy profile is
  similar to that predicted from experimental observations.
  Calculating free energies employing our simple method within TPS 
  provides significant advantages over such methods as umbrella sampling 
  since it is free from any applied external bias, accurate compared to 
  experimental measurements and has a reasonable computational cost.    
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Molecular dynamics (MD) simulations are increasingly being
used to calculate quantitative estimates of experimentally measurable 
kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
\cite{Karplus02NatStructMolBiol9p646,
Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
This has helped in proposing mechanisms and explanations for 
enzyme activity, fast high throughput screening of thousands of inhibitor 
molecules in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
directed evolution for designing enzymes with enhanced catalytic 
activities, \cite{Nunez06JPhysChemA110p463,Thyme09Nature461p1300,
Bloom09PNAS106p9995,Zoi16JAmChemSoc138p3403,Schafer19JAmChemSoc141p10431} etc.    
Straightforward MD simulations adequately sample 
the long-lived stable reactant and product states 
but the high energy barrier-crossing events that occur rarely and at long 
time scales are not easily accessible using this method alone.
To access such states at a higher frequency using MD simulations 
several enhanced sampling techniques 
have been developed in the past few decades. These include 
umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
steered MD, \cite{Park04JChemPhys120p5946}
milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,
Dellago98JChemPhys108p1964,Bolhuis21AdvTheory4p2000237}

TPS is an attractive method since it samples reactive trajectories
in which the rare barrier-crossing events are guaranteed to occur and
it does not require a priori knowledge of the reaction coordinate which can be 
complex for large biological systems.
TPS is based on Monte Carlo importance sampling in the 
space of trajectories. It uses the shooting and shifting algorithm 
to collect an ensemble of reactive pathways that are real dynamical 
trajectories generated without the application of any bias forces. \cite{dellago02AdvChemPhys123} 
Therefore the TPS ensemble can be used to obtain 
kinetic information such as the rate of the reaction as well as to elucidate 
the transition state, reaction coordinate and the reaction mechanism. 
TPS has been successfully applied to study problems such as enzyme catalysis, 
\cite{Basner05JAmChemSoc127p13822,Ramon07JPhysChemB111p5708,Hay12NatChem4p161,Schwartz09NatChemBiol5p551} 
and protein folding. \cite{Bolhuis03ProcNatlAcadSci100p12129,Juraszek12ChemPhys396p30}
